% Keine Kopf-/Fußzeile auf dieser Seite
\thispagestyle{empty}

%\vspace*{4cm}
\section*{Erkl\"arung zur Abschlussarbeit gem\"a\ss{} \S23 Abs. 7 APB der TU Darmstadt}\index{Erklärung}
\addcontentsline{toc}{chapter}{Erk\"arung zur Abschlussarbeit}

Hiermit versichere ich, Fabian Gabel, die vorliegende Master-Thesis ohne Hilfe Dritter und nur mit den angegebenen Quellen und Hilfsmitteln angefertigt zu haben. Alle Stellen, die Quellen entnommen wurden, sind als solche kenntlich gemacht worden. Diese Arbeit hat in gleicher oder \"ahnlicher Form noch keiner Pr\"ufungsbeh\"orde vorgelegen. 

Mir ist bekannt, dass im Falle eines Plagiats (\S38 Abs.2 APB) ein T\"auschungsversuch vorliegt, der dazu f\"uhrt, dass die Arbeit mit 5,0 bewertet und damit ein Pr\"ufungsversuch verbraucht wird. Abschlussarbeiten d\"urfen nur einmal wiederholt werden.

Bei der abgegebenen Thesis stimmen die schriftliche und die zur Archivierung eingereichte elektronische Fassung überein.

%\section*{Thesis Statement pursuant to \S23 paragraph 7 of APB TU Darmstadt}
\section*{English translation for information purposes only:\\
Thesis Statement pursuant to § 22 paragraph 7 and § 23 paragraph 7 of APB TU Darmstadt}

I herewith formally declare that I, Muster Mustermann, have written the submitted thesis independently pursuant to § 22 paragraph 7 of APB TU Darmstadt. I did not use any outside support except for the quoted literature and other sources mentioned in the paper. I clearly marked and separately listed all of the literature and all of the other sources which I employed when producing this academic work, either literally or in content. This thesis has not been handed in or published before in the same or similar form.

I am aware, that in case of an attempt at deception based on plagiarism (§38 Abs. 2 APB), the thesis would be graded with 5,0 and counted as one failed examination attempt. The thesis may only be repeated once.

In the submitted thesis the written copies and the electronic version for archiving are pursuant to § 23 paragraph 7 of APB identical in content.

%I herewith formally declare that I, Fabian Gabel, have written the submitted thesis independently. I did not use any outside support except for the quoted literature and other sources mentioned in the paper. I clearly marked and separately listed all of the literature and all of the other sources which I employed when producing this academic work, either literally or in content. This thesis has not been handed in or published before in the same or similar form.

%I am aware, that in case of an attempt at deception based on plagiarism (\S38 Abs. 2 APB), the thesis would be graded with 5,0 and counted as one failed examination attempt. The thesis may only be repeated once.

%In the submitted thesis the written copies and the electronic version for archiving are identical in content.
\vspace{20pt}


\noindent
Ort/Place, Datum/Date\vspace{30pt}

% Unterschrift (handgeschrieben)

\noindent
Unterschrift des Autors/Signature of author
